---
title: "MATRIX_dataprep"
author: "Mara DeMers"
date: "2026-09-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


2022 data - wheat and non-wheat ITS and 16S. ISDs removed (ASVs 1-3) but counts used to normalize count data. dada2 and lulu curation already complete. Metadata includes field collection data and relevant climate data.

Used BLAST to ISD sequences to mark for removal.  These have k__ISD in Kingdom field of taxonomy file (fungi). Use match_list_ISD.txt to remove from bacteria (preferred method).  Filter Mitochondria (Family) and Chloroplast (Order) taxa using phyloseq functions.

```{r, echo=FALSE}
library(ggplot2)
library(phyloseq)
library(vegan)
library(decontam)
library(microbiome)
library(DESeq2)
library(psadd)
```

Load files

```{r}
#load lulu output files
count_its <- read.csv("data/ASV_counts_lulu_f.csv", header = TRUE, row.names = 1, check.names = FALSE)
count_16s <- read.csv("data/ASV_counts_lulu_b.csv", header = TRUE, row.names = 1, check.names = FALSE)
tax_its <- as.matrix(read.csv("data/ASV_taxonomy_lulu_f.csv", header = TRUE, row.names = 1,
                              check.names = FALSE))
tax_16s <- as.matrix(read.csv("data/ASV_taxonomy_lulu_b.csv", header = TRUE, row.names = 1,
                              check.names = FALSE))
#load metadata
samdf_its <- read.csv("data/MATRIX22_ITS_samdf.csv", header = TRUE)
samdf_16s <- read.csv("data/MATRIX22_16S_samdf.csv", header = TRUE)
colnames(samdf_its)[1] <- "SampleID"
rownames(samdf_its) <- samdf_its$SampleID
colnames(samdf_16s)[1] <- "SampleID"
rownames(samdf_16s) <- samdf_16s$SampleID

samdf_its$LibraryGroup <- as.factor(samdf_its$LibraryGroup)
samdf_its$SeqGroup <- as.factor(samdf_its$SeqGroup)

samdf_16s$LibraryGroup <- as.factor(samdf_16s$LibraryGroup)
samdf_16s$SeqGroup <- as.factor(samdf_16s$SeqGroup)

#Define ps objects
ps_its <- phyloseq(otu_table(count_its, taxa_are_rows=T),
               sample_data(samdf_its), tax_table(tax_its))
ps_16s <- phyloseq(otu_table(count_16s, taxa_are_rows=T),
               sample_data(samdf_16s), tax_table(tax_16s))
```

Find index hopping threshold - largest ISD read count from PCR2 blank negative control samples

Other negative controls had ISDs added and so their presence does not indicate index hopping

```{r, echo=FALSE, eval = FALSE}
controlF <- phyloseq::subset_samples(ps_its, Collection_criteria == "PCR2_blank")

controlB <- phyloseq::subset_samples(ps_16s, Collection_criteria == "PCR2_blank")

ISDmatchF <- read.table("data/isd_match_list_f.txt") #BLAST hits to ISD sequences
ISDsF <- ISDmatchF$V1
allTaxaF = taxa_names(controlF)
ISDTaxaF = allTaxaF[(allTaxaF %in% ISDsF)]
controlF = prune_taxa(ISDTaxaF, controlF)
max(controlF@otu_table) #845 highest read count of ISD in PCR2 blank

ISDmatchB <- read.table("data/isd_match_list_b.txt") #BLAST hits to ISD sequences
ISDsB <- ISDmatchB$V1
allTaxaB = taxa_names(controlB)
ISDTaxaB = allTaxaB[(allTaxaB %in% ISDsB)]
controlB = prune_taxa(ISDTaxaB, controlB)
max(controlB@otu_table) #1245 highest read count of ISD in PCR2 blank
```

Function for changing NAs to the lowest assigned rank so the ASVs don't get killed by subset_taxa
Did not write this, got it from github under phyloseq issues

```{r, echo = FALSE, eval = FALSE}
tax.clean <- data.frame(row.names = row.names(tax_16s),
Kingdom = str_replace(tax_16s[,1], "D_0__",""),
Phylum = str_replace(tax_16s[,2], "D_1__",""),
Class = str_replace(tax_16s[,3], "D_2__",""),
Order = str_replace(tax_16s[,4], "D_3__",""),
Family = str_replace(tax_16s[,5], "D_4__",""),
Genus = str_replace(tax_16s[,6], "D_5__",""),
stringsAsFactors = FALSE)
tax.clean[is.na(tax.clean)] <- ""

for (i in 1:6){ tax.clean[,i] <- as.character(tax.clean[,i])}
####### Fill holes in the tax table
tax.clean[is.na(tax.clean)] <- ""
for (i in 1:nrow(tax.clean)){
  if (tax.clean[i,2] == ""){
kingdom <- paste("Kingdom_", tax.clean[i,1], sep = "")
tax.clean[i, 2:6] <- kingdom
} else if (tax.clean[i,3] == ""){
phylum <- paste("Phylum_", tax.clean[i,2], sep = "")
tax.clean[i, 3:6] <- phylum
} else if (tax.clean[i,4] == ""){
class <- paste("Class_", tax.clean[i,3], sep = "")
tax.clean[i, 4:6] <- class
} else if (tax.clean[i,5] == ""){
order <- paste("Order_", tax.clean[i,4], sep = "")
tax.clean[i, 5:6] <- order
} else if (tax.clean[i,6] == ""){
family <- paste("Family_", tax.clean[i,5], sep = "")
tax.clean[i, 6:7] <- family
}
}
tax_table(ps_16s) <- as.matrix(tax.clean)

##ITS
tax.clean <- data.frame(row.names = row.names(tax_its),
Kingdom = str_replace(tax_its[,1], "D_0__",""),
Phylum = str_replace(tax_its[,2], "D_1__",""),
Class = str_replace(tax_its[,3], "D_2__",""),
Order = str_replace(tax_its[,4], "D_3__",""),
Family = str_replace(tax_its[,5], "D_4__",""),
Genus = str_replace(tax_its[,6], "D_5__",""),
Species = str_replace(tax_its[,7], "D_6__",""),
stringsAsFactors = FALSE)
tax.clean[is.na(tax.clean)] <- ""

for (i in 1:7){ tax.clean[,i] <- as.character(tax.clean[,i])}
####### Fill holes in the tax table
tax.clean[is.na(tax.clean)] <- ""
for (i in 1:nrow(tax.clean)){
  if (tax.clean[i,2] == ""){
kingdom <- paste("Kingdom_", tax.clean[i,1], sep = "")
tax.clean[i, 2:7] <- kingdom
} else if (tax.clean[i,3] == ""){
phylum <- paste("Phylum_", tax.clean[i,2], sep = "")
tax.clean[i, 3:7] <- phylum
} else if (tax.clean[i,4] == ""){
class <- paste("Class_", tax.clean[i,3], sep = "")
tax.clean[i, 4:7] <- class
} else if (tax.clean[i,5] == ""){
order <- paste("Order_", tax.clean[i,4], sep = "")
tax.clean[i, 5:7] <- order
} else if (tax.clean[i,6] == ""){
family <- paste("Family_", tax.clean[i,5], sep = "")
tax.clean[i, 6:7] <- family
} else if (tax.clean[i,7] == ""){
tax.clean$Species[i] <- paste("Genus",tax.clean$Genus[i], sep = "_")
}
}
tax_table(ps_its) <- as.matrix(tax.clean)
```

Remove ISDs, non-target taxa

```{r, echo=FALSE, eval = FALSE}
#Remove non-fungi, non-bacteria, ISDs, NAs
allTaxaF = taxa_names(ps_its)
bioTaxaF = allTaxaF[!(allTaxaF %in% ISDsF)]
ps_its = prune_taxa(bioTaxaF, ps_its)

allTaxaB = taxa_names(ps_16s)
bioTaxaB = allTaxaB[!(allTaxaB %in% ISDsB)]
ps_16s = prune_taxa(bioTaxaB, ps_16s)

phyloseq::get_taxa_unique(ps_its, taxonomic.rank = "Kingdom")
ps_its <- phyloseq::subset_taxa(ps_its, Kingdom == "k__Fungi") 
ps_its <- phyloseq::prune_taxa(phyloseq::taxa_sums(ps_its) > 0, ps_its) #remove all-zeros
phyloseq::get_taxa_unique(ps_16s, taxonomic.rank = "Kingdom")
ps_16s <- phyloseq::subset_taxa(ps_16s, Kingdom == "Bacteria")
ps_16s <- phyloseq::subset_taxa(ps_16s, Order != "Chloroplast")
ps_16s <- phyloseq::subset_taxa(ps_16s, Family != "Mitochondria")
ps_16s <- phyloseq::subset_taxa(ps_16s, !Genus == "Buchnera")
ps_16s <- phyloseq::subset_taxa(ps_16s, !Genus == "Wolbachia")
ps_16s <- phyloseq::prune_taxa(phyloseq::taxa_sums(ps_16s) > 0, ps_16s) #remove all-zeros
```

Decontam - Wanted a more rigorous decontamination, so I am using the "either" method which calls contaminants identified by either frequency or prevalence methods.  No mercy just toss them out.

Decontam is run per sequencing group then samples are merged.

```{r, eval = FALSE}
#decontam frequency method requires positive numeric quantitation values
#Convert negative library concentration values to positive
vecf = sample_data(ps_its)$LibraryConc_ng.ul
vecf[vecf < 0.01] <- 0.01
sample_data(ps_its)$LibraryConc_ng.ul = vecf

vecb = sample_data(ps_16s)$LibraryConc_ng.ul
vecb[vecb < 0.01] <- 0.01
sample_data(ps_16s)$LibraryConc_ng.ul = vecb

#decontam step default thresholds
sample_data(ps_its)$is.neg <- sample_data(ps_its)$Source == "neg_ctl"
ps_its1 <- phyloseq::subset_samples(ps_its, SeqGroup == "1")
ps_its2 <- phyloseq::subset_samples(ps_its, SeqGroup == "2")
ps_its3 <- phyloseq::subset_samples(ps_its, SeqGroup == "3")
ps_its4 <- phyloseq::subset_samples(ps_its, SeqGroup == "4")
ps_its5 <- phyloseq::subset_samples(ps_its, SeqGroup == "5")
ps_its6 <- phyloseq::subset_samples(ps_its, SeqGroup == "6")
ps_its7 <- phyloseq::subset_samples(ps_its, SeqGroup == "7")
ps_its8 <- phyloseq::subset_samples(ps_its, SeqGroup == "8")
contamdf.its1 <- isContaminant(ps_its1, method="either", conc="LibraryConc_ng.ul", neg = "is.neg")
contamdf.its2 <- isContaminant(ps_its2, method="either", conc="LibraryConc_ng.ul", neg = "is.neg")
contamdf.its3 <- isContaminant(ps_its3, method="either", conc="LibraryConc_ng.ul", neg = "is.neg")
contamdf.its4 <- isContaminant(ps_its4, method="either", conc="LibraryConc_ng.ul", neg = "is.neg")
contamdf.its5 <- isContaminant(ps_its5, method="either", conc="LibraryConc_ng.ul", neg = "is.neg")
contamdf.its6 <- isContaminant(ps_its6, method="either", conc="LibraryConc_ng.ul", neg = "is.neg")
contamdf.its7 <- isContaminant(ps_its7, method="either", conc="LibraryConc_ng.ul", neg = "is.neg")
contamdf.its8 <- isContaminant(ps_its8, method="either", conc="LibraryConc_ng.ul", neg = "is.neg")

sample_data(ps_16s)$is.neg <- sample_data(ps_16s)$Source == "neg_ctl"
ps_16s1 <- phyloseq::subset_samples(ps_16s, SeqGroup == "1")
ps_16s2 <- phyloseq::subset_samples(ps_16s, SeqGroup == "2")
ps_16s3 <- phyloseq::subset_samples(ps_16s, SeqGroup == "3")
ps_16s4 <- phyloseq::subset_samples(ps_16s, SeqGroup == "4")
ps_16s5 <- phyloseq::subset_samples(ps_16s, SeqGroup == "5")
ps_16s6 <- phyloseq::subset_samples(ps_16s, SeqGroup == "6")
ps_16s7 <- phyloseq::subset_samples(ps_16s, SeqGroup == "7")
ps_16s8 <- phyloseq::subset_samples(ps_16s, SeqGroup == "8")
contamdf.16s1 <- isContaminant(ps_16s1, method="either", conc="LibraryConc_ng.ul", neg = "is.neg")
contamdf.16s2 <- isContaminant(ps_16s2, method="either", conc="LibraryConc_ng.ul", neg = "is.neg")
contamdf.16s3 <- isContaminant(ps_16s3, method="either", conc="LibraryConc_ng.ul", neg = "is.neg")
contamdf.16s4 <- isContaminant(ps_16s4, method="either", conc="LibraryConc_ng.ul", neg = "is.neg")
contamdf.16s5 <- isContaminant(ps_16s5, method="either", conc="LibraryConc_ng.ul", neg = "is.neg")
contamdf.16s6 <- isContaminant(ps_16s6, method="either", conc="LibraryConc_ng.ul", neg = "is.neg")
contamdf.16s7 <- isContaminant(ps_16s7, method="either", conc="LibraryConc_ng.ul", neg = "is.neg")
contamdf.16s8 <- isContaminant(ps_16s8, method="either", conc="LibraryConc_ng.ul", neg = "is.neg")

table(contamdf.its1$contaminant)  #8 TRUE
table(contamdf.its2$contaminant)  #17 TRUE
table(contamdf.its3$contaminant)  #15 TRUE
table(contamdf.its4$contaminant)  #34 TRUE
table(contamdf.its5$contaminant)  #5 TRUE
table(contamdf.its6$contaminant)  #12 TRUE
table(contamdf.its7$contaminant)  #33 TRUE
table(contamdf.its8$contaminant)  #12 TRUE
table(contamdf.16s1$contaminant)  #92 TRUE
table(contamdf.16s2$contaminant)  #156 TRUE
table(contamdf.16s3$contaminant)  #171 TRUE
table(contamdf.16s4$contaminant)  #80 TRUE
table(contamdf.16s5$contaminant)  #81 TRUE
table(contamdf.16s6$contaminant)  #76 TRUE
table(contamdf.16s7$contaminant)  #50 TRUE
table(contamdf.16s8$contaminant)  #16 TRUE

#Remove contaminants from phyloseq object
ps_nc_its1 <- prune_taxa(!contamdf.its1$contaminant, ps_its1)
ps_nc_its2 <- prune_taxa(!contamdf.its2$contaminant, ps_its2)
ps_nc_its3 <- prune_taxa(!contamdf.its3$contaminant, ps_its3)
ps_nc_its4 <- prune_taxa(!contamdf.its4$contaminant, ps_its4)
ps_nc_its5 <- prune_taxa(!contamdf.its5$contaminant, ps_its5)
ps_nc_its6 <- prune_taxa(!contamdf.its6$contaminant, ps_its6)
ps_nc_its7 <- prune_taxa(!contamdf.its7$contaminant, ps_its7)
ps_nc_its8 <- prune_taxa(!contamdf.its8$contaminant, ps_its8)
ps_nc_16s1 <- prune_taxa(!contamdf.16s1$contaminant, ps_16s1)
ps_nc_16s2 <- prune_taxa(!contamdf.16s2$contaminant, ps_16s2)
ps_nc_16s3 <- prune_taxa(!contamdf.16s3$contaminant, ps_16s3)
ps_nc_16s4 <- prune_taxa(!contamdf.16s4$contaminant, ps_16s4)
ps_nc_16s5 <- prune_taxa(!contamdf.16s5$contaminant, ps_16s5)
ps_nc_16s6 <- prune_taxa(!contamdf.16s6$contaminant, ps_16s6)
ps_nc_16s7 <- prune_taxa(!contamdf.16s7$contaminant, ps_16s7)
ps_nc_16s8 <- prune_taxa(!contamdf.16s8$contaminant, ps_16s8)
ps_nc_its <- phyloseq::merge_phyloseq(ps_nc_its1, ps_nc_its2, ps_nc_its3, ps_nc_its4, ps_nc_its5,
                                       ps_nc_its6, ps_nc_its7, ps_nc_its8)
ps_nc_16s <- phyloseq::merge_phyloseq(ps_nc_16s1, ps_nc_16s2, ps_nc_16s3, ps_nc_16s4, ps_nc_16s5,
                                       ps_nc_16s6, ps_nc_16s7, ps_nc_16s8)
ps_nc_its <- prune_samples(sample_data(ps_nc_its)$Source == "sample", ps_nc_its)
ps_nc_16s <- prune_samples(sample_data(ps_nc_16s)$Source == "sample", ps_nc_16s)

ps_nc_its <- phyloseq::prune_taxa(phyloseq::taxa_sums(ps_nc_its) > 0, ps_nc_its) #remove all-zeros
ps_nc_16s <- phyloseq::prune_taxa(phyloseq::taxa_sums(ps_nc_16s) > 0, ps_nc_16s)
#save(ps_nc_its, ps_nc_16s, file = "decontam.Rdata")

ASV_ITS <- phyloseq::otu_table(ps_nc_its)
#write.csv(ASV_ITS, "data/ASV_counts_decontam_f.csv")
ASV_16S <- phyloseq::otu_table(ps_nc_16s)
#write.csv(ASV_16S, "data/ASV_counts_decontam_b.csv")
TAX_ITS <- phyloseq::tax_table(ps_nc_its)
#write.csv(TAX_ITS, "data/ASV_taxonomy_decontam_f.csv")
TAX_16S <- phyloseq::tax_table(ps_nc_16s)
#write.csv(TAX_16S, "data/ASV_taxonomy_decontam_b.csv")
```

Remove observations below index-hopping threshold - determined by reads in negative control samples that should have nothing in them (vs. other negative controls with stuff in them) (845 fungi + 1245 bacteria)

```{r, echo=FALSE, eval = FALSE}
ps_nc_its@otu_table[ps_nc_its@otu_table < 845] <- 0
ps_nc_its <- phyloseq::prune_taxa(phyloseq::taxa_sums(ps_nc_its) > 0, ps_nc_its) #remove all-zeros
ps_nc_its <- phyloseq::prune_samples(phyloseq::sample_sums(ps_nc_its) > 0, ps_nc_its)
ps_nc_its # now only 3191 taxa, 2648 samples (from 2682 samples)

ps_nc_16s@otu_table[ps_nc_16s@otu_table < 1245] <- 0
ps_nc_16s <- phyloseq::prune_taxa(phyloseq::taxa_sums(ps_nc_16s) > 0, ps_nc_16s) #remove all-zeros
ps_nc_16s <- phyloseq::prune_samples(phyloseq::sample_sums(ps_nc_16s) > 0, ps_nc_16s)
ps_nc_16s # now only 7769 taxa, 2518 samples (from 2681 samples)

#save(ps_nc_its, ps_nc_16s, file = "indexhoppingremoved.RData")

ASV_ITS <- phyloseq::otu_table(ps_nc_its)
#write.csv(ASV_ITS, "data/ASV_counts_IHThresh_f.csv")
ASV_16S <- phyloseq::otu_table(ps_nc_16s)
#write.csv(ASV_16S, "datab/ASV_counts_IHThresh_b.csv")
TAX_ITS <- phyloseq::tax_table(ps_nc_its)
#write.csv(TAX_ITS, "data/ASV_taxonomy_IHThresh_f.csv")
TAX_16S <- phyloseq::tax_table(ps_nc_16s)
#write.csv(TAX_16S, "datab/ASV_taxonomy_IHThresh_b.csv")
```

Normalize counts from ISD data (I used to do this in Excel but then I got fancy)

```{r, echo=FALSE, eval = FALSE}
ps_its <- prune_samples(sample_data(ps_its)$Source == "sample", ps_its)
ps_16s <- prune_samples(sample_data(ps_16s)$Source == "sample", ps_16s)

#we already got rid of ISDs so go get them back
normISDF <- phyloseq::prune_taxa(c("ASV_00001", "ASV_00003", "ASV_00017"), ps_its)
normISDF <- phyloseq::prune_samples(phyloseq::sample_names(ps_nc_its), normISDF)
normISDF
normISDB <- phyloseq::prune_taxa(c("ASV_000001", "ASV_000002", "ASV_000003"), ps_16s)
normISDB <- phyloseq::prune_samples(phyloseq::sample_names(ps_nc_16s), normISDB)
normISDB

samplesumF <- phyloseq::sample_sums(normISDF)
samplesumB <- phyloseq::sample_sums(normISDB)

#normalizing
normbyF <- samplesumF/3
normbyF <- as.vector(normbyF)
asv_its <- t(ps2m_otu(ps_nc_its))

normbyB <- samplesumB/3
normbyB <- as.vector(normbyB)
asv_16s <- t(ps2m_otu(ps_nc_16s))

#Define ISD copies per reaction - ITS
#ISD_copies_per_rxn = 585
#template_ug_per_rxn = 2.5
ISD_copies_per_ug_template = 585/2.5
#abs = mat / vec[col(mat)] * ISD_copies_per_ug_template
asv_its_norm <- sweep(asv_its, 2, normbyF, "/")
asv_its_norm <- asv_its_norm * ISD_copies_per_ug_template

#Define ISD copies per reaction - 16S
#ISD_copies_per_rxn = 18
#template_ug_per_rxn = 6
ISD_copies_per_ug_template = 18/6
#abs = mat / vec[col(mat)] * ISD_copies_per_ug_template
asv_16s_norm <- sweep(asv_16s, 2, normbyB, "/")
asv_16s_norm <- asv_16s_norm * ISD_copies_per_ug_template

ps_its_n <- phyloseq(otu_table(asv_its_norm, taxa_are_rows=T),
                     sample_data(samdf_its), tax_table(tax_its))
ps_16s_n <- phyloseq(otu_table(asv_16s_norm, taxa_are_rows=T),
                     sample_data(samdf_16s), tax_table(tax_16s))

ps_its_n <- phyloseq::prune_samples(phyloseq::sample_sums(ps_its_n) > 0, ps_its_n)
ps_its_n

ps_16s_n <- phyloseq::prune_samples(phyloseq::sample_sums(ps_16s_n) > 0, ps_16s_n)
ps_16s_n

#save(ps_its_n, ps_16s_n, file = "normalized.RData")

ASV_ITS_N <- phyloseq::otu_table(ps_its_n)
write.csv(ASV_ITS_N, "data/ASV_counts_normalized_f.csv")
ASV_16S_N <- phyloseq::otu_table(ps_16s_n)
write.csv(ASV_16S_N, "data/ASV_counts_normalized_b.csv")
TAX_ITS_N <- phyloseq::tax_table(ps_its_n)
write.csv(TAX_ITS_N, "data/ASV_taxonomy_normalized_f.csv")
TAX_16S_N <- phyloseq::tax_table(ps_16s_n)
write.csv(TAX_16S_N, "data/ASV_taxonomy_normalized_b.csv")
```

Transforming normalized abundances - CLR making a sunburst pattern isn't a good thing!  If not transforming is best, just do that.

OH NO this takes so long with large data sets.  Do this once and save it and never do it again.

```{r, echo = FALSE, eval = FALSE}
#Trying VST transform then comparing to raw counts and normalized counts (mean-variance)
ps_ds_its <- phyloseq::phyloseq_to_deseq2(ps_its_n, ~Month+Source_plant+SeqGroup)
gm_mean = function(x, na.rm=TRUE){ exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))}
geoMeans = apply(counts(ps_ds_its), 1, gm_mean)
ps_ds_its = estimateSizeFactors(ps_ds_its, type="ratio", geoMeans = geoMeans)
#ps_ds_its = DESeq2::estimateSizeFactors(ps_ds_its)
ps_ds_its = DESeq2::estimateDispersions(ps_ds_its, fitType = "local")
DESeq2::plotDispEsts(ps_ds_its)
ps_vst_its <- ps_its_n
vstF <- DESeq2::getVarianceStabilizedData(ps_ds_its)
row.names(vstF) #make sure taxa are rows
phyloseq::otu_table(ps_vst_its) <- phyloseq::otu_table(vstF, taxa_are_rows = TRUE)
ps_vst_its

ps_ds_16s <- phyloseq::phyloseq_to_deseq2(ps_16s_n, ~Month+Source_plant+SeqGroup)
geoMeans = apply(counts(ps_ds_16s), 1, gm_mean)
ps_ds_16s = estimateSizeFactors(ps_ds_16s, type="ratio", geoMeans = geoMeans)
#ps_ds_16s = DESeq2::estimateSizeFactors(ps_ds_16s)
ps_ds_16s = DESeq2::estimateDispersions(ps_ds_16s, fitType = "local")
DESeq2::plotDispEsts(ps_ds_16s)
ps_vst_16s <- ps_nc_16s
vstB <- DESeq2::getVarianceStabilizedData(ps_ds_16s)
row.names(vstB) #make sure taxa are rows
phyloseq::otu_table(ps_vst_16s) <- phyloseq::otu_table(vstB, taxa_are_rows = TRUE)
ps_vst_16s
#for some reason deseq insists there are NAs but anyNA says there are none. nothing to be done

#alternative - CLR transform
ps_clr_its <- microbiome::transform(ps_its_n, transform = "clr")
ps_clr_16s <- microbiome::transform(ps_16s_n, transform = "clr")

p1 <- plot_mv(ps_nc_its, log=FALSE, title = "Raw - Fungi")
p2 <- plot_mv(ps_its_n, log=FALSE, title = "Normalized - Fungi")
p3 <- plot_mv(ps_vst_its, log=FALSE, title = "VST - Fungi")
p4 <- plot_mv(ps_clr_its, log=FALSE, title = "CLR - Fungi")
p5 <- plot_mv(ps_nc_16s, log=FALSE, title = "Raw - Bacteria")
p6 <- plot_mv(ps_16s_n, log=FALSE, title = "Normalized - Bacteria")
plot_mv(ps_vst_16s, log=FALSE, title = "VST - Bacteria")
p7 <- plot_mv(ps_clr_16s, log=FALSE, title = "CLR - Bacteria")
ggpubr::ggarrange(p1, p2, p3, p4,
                  ncol = 2, nrow = 2, common.legend = TRUE)
ggpubr::ggarrange(p5, p6, p7,
                  ncol = 2, nrow = 2, common.legend = TRUE)

#save(ps_vst_its, ps_clr_its, ps_vst_16s, ps_clr_16s, file = "transformednormalcounts.RData")

#since VST isn't working for 16s and CLR looks okay for both I'll use that
```

Subset to only samples from March or April (early or late season sampling) - we've got a lot more samples than that but you don't need them if you're just copying what we did in the paper

```{r, echo=FALSE}
ps_nc_its <- phyloseq::subset_samples(ps_nc_its, Month %in% c("April", "March"))
ps_clr_its <- phyloseq::subset_samples(ps_clr_its, Month %in% c("April", "March"))
ps_its_n <- phyloseq::subset_samples(ps_its_n, Month %in% c("April", "March"))
ps_nc_16s <- phyloseq::subset_samples(ps_nc_16s, Month %in% c("April", "March"))
ps_clr_16s <- phyloseq::subset_samples(ps_clr_16s, Month %in% c("April", "March"))
ps_16s_n <- phyloseq::subset_samples(ps_16s_n, Month %in% c("April", "March"))

save(ps_nc_its, ps_clr_its, ps_its_n, ps_nc_16s, ps_clr_16s, ps_16s_n, file = "MarchAprilOnly.RData")
```